---
title: "Book Program Analysis"
author: "Joshua Goldberg"
date: "9/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Summary
During this project, I investigated book sales and reviews over time using sentiment analysis to understand how the user submitted reviews correlate with the purchase totals for each book for this new program.


## Goal
Throughout the analysis, the goal was to analyze whether or not this new program was successful and increasing sales and improving review quality.


## Introduction
When exploring the data that was provided for this project, there were six data science books that I also provided with user feedback reviews on. Using the reviews, I was able to probe for trends and insights by looking at purchase rates over time and to see which other factors underwent significant total changes after given dates.


## Findings
From this analysis, it appears that overall, review quality decreased after the inception of the new program, and when we examine these differences for the two customer groups, Business increase while Individuals decrease in quality. Similarly, the sales after the new program 
decrease overall, and for each customer groups, the totals change as does review quality, where Business sales increase while Individual sales decrease.

# Conclusion
Overall, looking into this data provided a great opportunity to leverage skills in mapping and vectorization, which can be a powerful data science skill to apply when performing function calculations on vectors of data, as I conducted here with book sales and sentiment reviews.


## Analysis
```{r analysis}

#load in data
book_program <- read_csv("C:/Users/Goldbej6/Desktop/Learning/R/Book_Program_Project/sales2019.csv")


#quick df analysis
dim(book_program) ##5000 rows and 5 columns in dataset
glimpse(book_program)

#Column Names
vector_cols <- colnames(book_program)
vector_cols 
#the column names represent: (1) The date when a user submitted a book review, (2) The specific text response from their review, (3) The title of the book, (4) The total count of books that this person purchased by each person, (4) The customer type of each person


#Datatypes
col_types <- for (c in colnames(book_program)) {
  paste0("The Data Type of the '", c, "' column is: ",typeof(book_program[[c]])) %>% 
  print
}
  
col_types


#Unique Values
for (c in vector_cols) {
  print("Unique column values:")
  print(c)
  print(unique(book_program[[c]]))
  print(" ")
}


#Locating and Removing Missing Values
#The total_purchased column appears to have missing values
for (c in colnames(book_program)) {
  paste0(c,
         " column, This many rows have missing data: ",
         is.na(book_program[c]) %>% sum) %>% 
         print
}

print("We can see that two columns have missing data, the user_submitted_review column, with around 885 missing data rows, and the total_purchased column, with 718 missing data values")

book_program <- book_program %>% 
  filter(!(is.na(user_submitted_review)))

dim(book_program) # 585 rows were removed from this operation
      
tot_purchased_avg <- book_program %>% 
  filter(!is.na(total_purchased)) %>%
  pull(total_purchased) %>% 
  mean
        
# Mean imputation
book_program <- book_program %>% 
  mutate(
    total_purchased_new = if_else(is.na(total_purchased),
                                  tot_purchased_avg,
                                  total_purchased)
  )


# Adding in Sentiment Assessment
#Unique column values
book_program %>% 
  pull(user_submitted_review) %>% 
  unique


# Function to evaluate Sentiment
positive_sentiment <- function(review) {
  review_status = case_when(
    str_detect(review, "ok") ~ TRUE,
    str_detect(review, "awesome") ~ TRUE,
    str_detect(review, "a lot") ~ TRUE,
    str_detect(review, "not") ~ FALSE,
    str_detect(review, "hated") ~ FALSE,
    TRUE ~ FALSE
  )
}

# Adding new column to dataframe for Sentiment check
book_program <- book_program %>% 
  mutate(
    positive_sentiment_indicator = unlist(map(user_submitted_review, positive_sentiment))
  )

# Date/Time Conversions
library(lubridate)
book_program <- book_program %>% 
  mutate(
    Date_Time = lubridate::mdy(date)
  )


# Comparing effectivness of new Program
book_program <- book_program %>% 
  mutate(
    date_comparison = if_else(mdy(date) < mdy("07-01-2019"), "Before", "After")
  )


purch_comp <- book_program %>% 
  group_by(date_comparison) %>% 
  summarize(
    books_purchased = sum(total_purchased_new)
  )

print("The program was unfortunately not effective in terms of increasing the number of books sold")


# Additional group divisions
purch_comp <- book_program %>% 
  group_by(date_comparison, customer_type) %>% 
  summarize(
    books_purchased = sum(total_purchased_new)
  ) %>% 
  arrange(customer_type, date_comparison)


view(purch_comp)
print("It looks like the effectiveness of the program does differ somewhat based on the groupings - Businesses did do slightly better than did inidividuals, with a slight increased after the program.")


# Comparing Sentiment reviews before and after Date of Interest
reviews_comp <- book_program %>% 
  group_by(date_comparison) %>% 
  summarize(
    positive_review_count = sum(positive_sentiment_indicator)
  ) %>% 
  arrange(-positive_review_count)




view(reviews_comp)
print("The number of positive reviews actually slightly decreased after July 1, 2019. So, review sentiment got worse after the program was created.")


library(ggplot2)
```

## Comparing Book Profit

You can also embed plots, for example:

```{r pressure, echo=TRUE}
qplot(x = date_comparison,
      y = books_purchased,
      color = customer_type,
      data = purch_comp)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
